High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-10-20 Draft_V4046Sgr_mnras.tex
Update Draft_V4046Sgr_mnras.tex

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the {\tt uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a $\chi^2$ model image with $\lambda$ = 0.1 for regularization. The resulting {\tt uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at $\sim$35\,au into a shallow tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between
$\sim$10 and 17\,au.

\begin{figure*}
  \includegraphics[width=\textwidth]{hot_two_E.pdf}
  \caption{Comparison of observations and simulated images at 1.65\,$\micron$ and 1.3 mm continuum of the circumbinary disc orbiting V4046\,Sgr. From top to bottom: DPI image and ALMA Band 6 observations; simulated images of the parametric model. Top left panel: SPHERE-IRDIS \textit{H}-band image with a yellow filled circle that illustrate the N\_ALC\_YJH\_S coronagraph (inner working angle $\sim0\farcs12$, or $\sim$8.6\,au at 72.4\,pc). Top right panel: 1.3\,mm continuum {\tt uvmem} model image. The yellow ellipse shows the size of the natural-weighted beam ($ 0\farcs077 \, \times \, 0\farcs051$), and the small orange ellipse shows an estimated {\tt uvmem} beam size ($\sim0\farcs026 \, \times \, \sim0\farcs017$). The inset zooms into the central emission, and the green star marks the star positions. Bottom left panel: 1.65\,$\upmu m$ simulated image of the parametric model. Bottom right panel: 1.3\,mm simulated image of the parametric model. The orange ellipse shows the size of the beam: $0\farcs026 \, \times \, 0\farcs017$. For all the images in the figure the colour scale is linear.}
  \label{fig:two}
\end{figure*}

Ring13 is surprisingly narrow and seems to be off-centre relative to the GAIA stellar position (at the origin of coordinates in Fig. \ref{fig:two}). We determined the ring's centre and orientation by minimising the dispersion of the disc radial profiles, from 6\,au to 19\,au, and thus obtained a PA of 74.64$\pm$0.08\,$\degr$, an inclination of 33.90$\pm$0.04\,$\degr$ and a centre at $\Delta \alpha = 9\pm0.05$\,mas $\Delta \delta = 0.06\pm0.04$\,mas, relative to the stars.

@@

Interestingly, the ALMA image also detects faint 1.3\,mm continuum emission near the stellar positions (See the inset in Fig. \ref{fig:two}). Since this faint central emission is offset from the star, at the origin of coordinates in the figure, it is probably due to large dust grains. This dust structure is at a distance of only $\sim0\farcs02$ from the binary system, and may match the small grain ring called Ring6 introduced in the RT model to match the SED (see Sec.\,\ref{sec:model} below). If so, then given that the binary phase at the time of the observations is PA=347\,$\degr$, so that the secondary lies roughly northwards of the primary, it may be that the northern side of Ring6 is shadowed, resulting in a seemingly lopsided geometry.

The scattered-light image shown in Fig. \ref{fig:two} also shows a double ring structure in the micron-
sized dust distribution. The observed morphology presents an inner cavity of $\sim$10\,au in radius and two rings located at 14.10$\pm$0.01, coincident with Ring13, and 24.62$\pm$0.08\,au, coincident with Ring24, with a small gap between them at $\sim$20\,au. The observed second ring matches the inner wall of the 1.3\,mm continuum emission outer ring \citep{Ru_z_Rodr_guez_2019}. Two other important features that are present in the image are: the brightness asymmetry and the shadows projected on the disc produced by the close binary system as they eclipse each other, described in detail by \citet{dOrazi}.

@@

@@

For the representation of the stars, we used two Kurucz photospheres models \citep{1979ApJS...40....1K, 1997A&A...318..841C} with T$_{\mathrm{eff},1} =$ 4350\,K, R$_{*,1} =$ 1.064\,R$_{\sun}$, M$_{*,1} =$ 0.90\,M$_{\sun}$ and T$_{\mathrm{eff},2} =$ 4060\,K, R$_{*,2} =$ 1.033\,R$_{\sun}$, M$_{*,2} =$ 0.85\,M$_{\sun}$ respectively and with an accretion rate of log$\,\dot{\mathrm{M}} = -$9.3\,Myr$^{-1}$ for both cases \citep{10.1111/j.1365-2966.2011.19366.x}. The stars were placed at a  mutual separation of 0.041\,au, so that their centre of mass coincides with the origin,  and oriented at a PA of 250\,$\degr$, so that the secondary lies to the NWW from the primary, thus casting the same shadow as observed by d'Orazzi***  rotation, relative to the semi-major axis of the disk, to match the shadows reported by \citet{dOrazi}.

@@

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. ***explicar que es varphi y como se transforma en psi**** Our model uses the same values, with two different flaring indeces $\psi$. The separation between the two values was set at $r$ = 19\,au with $\psi=0.2$ for  Ring6 and Ring13, and $\psi=0.5$ for Ring24. The scale height is  $H_\circ = 0.045$ at $r_0 = 19.5$\,au.

@@

The observed near-far asymmetry in the DPI image is suggestive of a strongly forward-scattering phase function. In order to reproduce a similar  asymmetry, we required much larger grains that typically used in the RT modeling of  such near-IR data \citep[e.g.][]{2018MNRAS.477.5104C}.

The simulated DPI image at 1.65\,$\micron$ in Fig.\,\ref{fig:two} was obtained with the scattering matrix calculated by the {\tt makeopac.py} script provided in the RADMC3D package. We used a  Gaussian size distribution centred at 0.4\,$\micron$, and smeared  out by 30\% in both directions, and using 20 grain size bins within  that range.

@@

On the other hand, in order to reproduce the double-ringed  morphology of the millimetre continuum emission,  we required a model with   a wide central cavity ($r = 11.8$\,au), a narrow 3.6\,au-wide inner ring (which is Ring13) followed by a .2\,au gap  that separates the  outer ring (Ring24), which starts at  22\,au and reaches peak  densities  at 27.7\,au. This last ring has an observed break between 30.2 and 36.7\,au, which is reproduced in the model with the a discontinuity in the exponent $\gamma$, along with the sum of the decaying exponentials. We also truncated  the model large-grain density at  $\sim$55\,au.

@@

Our parametric model is fairly successful in reproducing the available data. The simulated images and the SED of the model are shown in Fig. \ref{fig:two} and Fig. \ref{fig:SED} respectively.

@@

The simulated 1.3\,mm continuum image  reproduces the two observed rings: the faint Ring13 and the  brighter Ring24. As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig. \ref{fig:radprofiles}), we can assume that the model provides an approximation of the disc structure, including the dimensions of  Ring13. Accordingly, taking the parametric model values for Ring13 we have that it has a radius of  13.6\,au,  a  width of 3.6\,au, and a  scale height ranging from 0.24 to 0.32\,au. The total dust mass of Ring13 is  about 3\,M$_{\earth}$ in the model.  In turn, in our model Ring24 reaches peak intensity at $\sim$30\,au, then  breaks  at $\sim$35\,au, as in the observations, and has a total dust  mass of $\sim$43\,M$_{\earth}$.

****DISCUTIR EL ANCHO OBSERVADO 3.87 VS. ANCHO MODELO 3.6 ****

@@

The observed SED was collected from data  in \citet{Jensen_97}, in \textsc{vizier} ***FOR WHICH BANDS?***  and \textsc{diana} ***idem***, and also using spectrometry data from an archival \textit{Spitzer} IRS spectrum.   ***THIS SHOULD GO IN SEC. OBS****

>>>>

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the {\tt uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a $\chi^2$ model image with $\lambda$ = 0.1 for regularization. The resulting {\tt uvmem} image shown in Fig. \ref{fig:images_vs_simulated} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at $\sim$35\,au into a shallow tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between
$\sim$10 and 17\,au.

\begin{figure*}
  \includegraphics[width=\textwidth]{hot_two_E.pdf}
  \caption{Comparison of observations and simulated images at 1.65\,$\micron$ and 1.3 mm continuum of the circumbinary disc orbiting V4046\,Sgr. From top to bottom: DPI image and ALMA Band 6 observations; simulated images of the parametric model. Top left panel: SPHERE-IRDIS \textit{H}-band image with a yellow filled circle that illustrate the N\_ALC\_YJH\_S coronagraph (inner working angle $\sim0\farcs12$, or $\sim$8.6\,au at 72.4\,pc). Top right panel: 1.3\,mm continuum {\tt uvmem} model image. The yellow ellipse shows the size of the natural-weighted beam ($ 0\farcs077 \, \times \, 0\farcs051$), and the small orange ellipse shows an estimated {\tt uvmem} beam size ($\sim0\farcs026 \, \times \, \sim0\farcs017$). The inset zooms into the central emission, and the green star marks the star positions. Bottom left panel: 1.65\,$\upmu m$ simulated image of the parametric model. Bottom right panel: 1.3\,mm simulated image of the parametric model. The orange ellipse shows the size of the beam: $0\farcs026 \, \times \, 0\farcs017$. For all the images in the figure the colour scale is linear.}
  \label{fig:images_vs_simulated}
\end{figure*}

Ring13 is surprisingly narrow and seems to be off-centre relative to the GAIA stellar position (at the origin of coordinates in Fig. \ref{fig:images_vs_simulated}). We determined the ring's centre and orientation by minimising the dispersion of the disc radial profiles, from 6\,au to 19\,au, and thus obtained a PA of 74.64$\pm$0.08\,$\degr$, an inclination of 33.90$\pm$0.04\,$\degr$ and a centre at $\Delta \alpha = 9\pm0.05$\,mas $\Delta \delta = 0.06\pm0.04$\,mas, relative to the stars.

@@

Interestingly, the ALMA image also detects faint 1.3\,mm continuum emission near the stellar positions (See the inset in Fig. \ref{fig:images_vs_simulated}). Since this faint central emission is offset from the star, at the origin of coordinates in the figure, it is probably due to large dust grains. This dust structure is at a distance of only $\sim0\farcs02$ from the binary system, and may match the small grain ring called Ring6 introduced in the RT model to match the SED (see Sec.\,\ref{sec:model} below). If so, then given that the binary phase at the time of the observations is PA=347\,$\degr$, so that the secondary lies roughly northwards of the primary, it may be that the northern side of Ring6 is shadowed, resulting in a seemingly lopsided geometry.

The scattered-light image shown in Fig. \ref{fig:images_vs_simulated} also shows a double ring structure in the micron-
sized dust distribution. The observed morphology presents an inner cavity of $\sim$10\,au in radius and two rings located at 14.10$\pm$0.01, coincident with Ring13, and 24.62$\pm$0.08\,au, coincident with Ring24, with a small gap between them at $\sim$20\,au. The observed second ring matches the inner wall of the 1.3\,mm continuum emission outer ring \citep{Ru_z_Rodr_guez_2019}. Two other important features that are present in the image are: the brightness asymmetry and the shadows projected on the disc produced by the close binary system as they eclipse each other, described in detail by \citet{dOrazi}.

@@

The observed SED was collected from data  in \citet{Jensen_97}, in \textsc{vizier} and \textsc{diana}, and also using spectrometry data from an archival \textit{Spitzer} IRS spectrum (See Fig.\ref{fig:SED}.). ***

@@

For the representation of the stars, we used two Kurucz photospheres models \citep{1979ApJS...40....1K, 1997A&A...318..841C} with T$_{\mathrm{eff},1} =$ 4350\,K, R$_{*,1} =$ 1.064\,R$_{\sun}$, M$_{*,1} =$ 0.90\,M$_{\sun}$ and T$_{\mathrm{eff},2} =$ 4060\,K, R$_{*,2} =$ 1.033\,R$_{\sun}$, M$_{*,2} =$ 0.85\,M$_{\sun}$ respectively and with an accretion rate of log$\,\dot{\mathrm{M}} = -$9.3\,Myr$^{-1}$ for both cases \citep{10.1111/j.1365-2966.2011.19366.x}. The stars were placed at a  mutual separation of 0.041\,au, so that their centre of mass coincides with the origin,  and oriented at a PA of 250\,$\degr$, so that the secondary lies to the NWW from the primary, thus casting the same shadow as observed by \citet{dOrazi}.

@@

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. ***explicar que es varphi y como se transforma en psi*** Our model uses the same values, with two different flaring indeces $\psi$. The separation between the two values was set at $r$ = 19\,au with $\psi=0.2$ for  Ring6 and Ring13, and $\psi=0.5$ for Ring24. The scale height is  $H_\circ = 0.045$ at $r_0 = 19.5$\,au.

@@

The observed near-far asymmetry in the DPI image is suggestive of a strongly forward-scattering phase function. In order to reproduce a similar  asymmetry, we required much larger grains that typically used in the RT modelling of  such near-IR data \citep[e.g.][]{2018MNRAS.477.5104C}.

The simulated DPI image at 1.65\,$\micron$ in Fig.\,\ref{fig:images_vs_simulated} was obtained with the scattering matrix calculated by the {\tt makeopac.py} script provided in the RADMC3D package. We used a  Gaussian size distribution centred at 0.4\,$\micron$, and smeared  out by 30\% in both directions, and using 20 grain size bins within  that range.

@@

On the other hand, in order to reproduce the double-ringed  morphology of the millimetre continuum emission,  we required a model with   a wide central cavity ($r = 11.7$\,au), a narrow 3.7\,au-wide inner ring (which is Ring13) followed by a .2\,au gap  that separates the  outer ring (Ring24), which starts at  22\,au and reaches peak  densities  at 27.7\,au. This last ring has an observed break between 30.2 and 36.7\,au, which is reproduced in the model with the a discontinuity in the exponent $\gamma$, along with the sum of the decaying exponentials. We also truncated  the model large-grain density at  $\sim$55\,au.

@@

Our parametric model is fairly successful in reproducing the available data. The simulated images and the SED of the model are shown in Fig. \ref{fig:images_vs_simulated} and Fig. \ref{fig:SED} respectively.

@@

The simulated 1.3\,mm continuum image  reproduces the two observed rings: the faint Ring13 and the  brighter Ring24. As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig. \ref{fig:radprofiles}), we can assume that the model provides an approximation of the disc structure, including the dimensions of  Ring13. Accordingly, taking the parametric model values for Ring13 we have that it has a radius of  13.6\,au,  a  width of 3.7\,au, and a  scale height ranging from 0.24 to 0.32\,au. The total dust mass of Ring13 is  about 3\,M$_{\earth}$ in the model.  In turn, in our model Ring24 reaches peak intensity at $\sim$30\,au, then  breaks  at $\sim$35\,au, as in the observations, and has a total dust  mass of $\sim$43\,M$_{\earth}$.

***DISCUTIR EL ANCHO OBSERVADO (3.87) VS. ANCHO MODELO (3.7) ***

@@
